---
title: "Transpop"
output: html_notebook
---

As an undergraduate researcher, I’ve had the privilege of exploring and familiarizing myself with a few public-use datasets and assisting my professors with various parts of the research process. One data set I’ve worked with several times is **“TransPop,”** a nationally representative survey on transgender health which includes data from 274 transgender and gender nonconforming (TGNC) adults as well as and 1,162 cisgender adults.

The recent attacks on TGNC rights have been on my mind a lot, and putting my energy into research helps me feel a bit more in control. When I got this assignment, I knew I wanted to look at TransPop again because I knew there was so much more to investigate. What I didn’t expect was the new banner which greeted me at the top of the research repository:

> On March 31, 2025, as a sponsor of this project, NIH requested that the following language be added to this website: This repository is under review for potential modification in compliance with Administration directives.

I cannot find the words to adequately express how angry I am about the politicization of parts of the human experience as beautiful as transness and the queer community. In lieu of an eloquent statement, I’d like to offer some hope.

**Project Background + Research Aims**

For this brief project, I decided to investigate the impact of stressors specific to the trans community on the social well-being and satisfaction with life of TGNC adults. By identifying patterns in the data, I endeavor to uncover what makes a TGNC person more likely to thrive, and thus, how we can help transgender lives flourish.

***Question: Which experiences and stressors (or the lack thereof) are most pertient to the social well-being and satisfaction with life of TGNC adults?***

**Procedure**

I chose to utilize a conditional inference tree model for my analysis. Conditional inference trees offer a robust method for regression analysis in decision trees and work well with multivariate analysis.

13 variables were included as potential predictors in the conditional inference tree models. Four total models were fitted; one tree with only the 4 demographic variables included as predictors for both outcome measures, and one tree with the 9 remaining variables included as predictors for both outcome measures.

For detailed information on the measures and survey items as well as full citations for the subscales included in this paper, please see the official methods and technical notes document for the TransPop survey (Krueger et al., 2020).

**Demographics**

Age, race, gender-identity, and sexual minority identity were included as demographic variables. Age was measured in years. For race, the variable in which responses were recoded as White, Black,Latino/Hispanic/Spanish origin, Multirace, and “other” was used for analysis (race_recode_cat5). Since the data from cisgender adults is not included in analysis for this paper, gender identity includes trans man (FTM), trans woman (MTF), or transgender gender nonbinary. Sexual minority identity is a binary label, with 0 representing hetereosexual and 1 representing any sexual minority identity labels (homosexual, bisexual etc).

**Measures**

I chose social well-being and satisfaction with life as my two outcome measures. Only complete cases were included for these composite scores.

**Social Well-Being**\
This 15-item scale (Keyes, 1998) evaluates individuals' perceived social wellness, including their sense of community belonging, comfort, and societal contribution. Participants rate statements like *"I don’t feel I belong to anything I’d call a community"* (reverse-coded) and *"My community is a source of comfort"* on a 7-point Likert scale (1 = *strongly disagree*, 7 = *strongly agree*). Eight items are reverse-coded before calculating a mean score (range: 1–7), with higher scores reflecting stronger social well-being—greater feelings of connection, support, and purpose in society.

**Satisfaction with Life Scale (SWLS)**\
The SWLS (Diener et al., 1985) measures global life satisfaction through five items (e.g., *"In most ways my life is close to ideal"* and *"I am satisfied with life"*), rated on a 7-point Likert scale (1 = *strongly disagree*, 7 = *strongly agree*). A mean score (range: 1–7) is computed, where higher scores indicate greater life satisfaction—more positive judgments about one’s life overall.

Each of the 9 variables included as predictors in this analysis are composite scores calculated from empirically validated scales included as survey items in the original questionnaire. Missing data was replaced with predicted mean matching imputation, with both imputed and non-imputed versions of the composite scores available for analysis within the dataset. I chose to use the imputed composite scores as predictors in my analysis. The predictors included were categorized as identity, healthcare access & utilization, or stressors.

**Identity:**

**Community Connectedness**: A 5-item scale (1–5) assessing transgender community affiliation; higher mean scores indicate stronger connectedness. 

**Non-Affirmation of Gender Identity**: A 6-item scale (1–5) measuring perceived lack of understanding/acceptance; higher mean scores indicate greater non-affirmation. 

**Healthcare access & utilization:**

**Healthcare Stereotype Threat**: A 4-item scale (1–5) assessing fear of being judged by providers due to LGBT identity; higher mean scores indicate greater worry. 

**Stressors:**

**Internalized Transphobia**: A 6-item scale (1–5) measuring self-directed stigma; higher mean scores indicate stronger internalized transphobia. 

**Everyday Discrimination**: A 9-item scale (1–4, reverse-coded) assessing chronic discrimination; higher mean scores indicate more frequent experiences. 

**Childhood Gender Conformity**: A 4-item scale (1–5) categorizing participants (based on sex at birth) into deciles of gender nonconformity (1 = most GNC, 3 = least GNC). 

**Gender Identity Non-Disclosure**: A 5-item scale (1–5) measuring avoidance of disclosing gender identity; higher mean scores indicate greater concealment efforts. 

**Negative Expectations of the Future**: A 9-item scale (1–5) assessing anticipated rejection due to gender identity; higher mean scores indicate more pessimistic expectations. 

**Social Support**: A 12-item scale (1–7) measuring perceived support; higher mean scores indicate stronger social support. 

```{r packages, include=FALSE}
library(partykit)
library(tidyverse)
```

```{r load data, include=FALSE}
transpop <- read.delim("37938-0005-Data.tsv")
```

```{r cleaning, include=FALSE}
transonly <- transpop |>
  filter(GENDER_IDENTITY > 2) |>
  mutate(GENDER_IDENTITY = as.factor(GENDER_IDENTITY)) |>
  select(
    STUDYID,
    WEIGHT_TRANSPOP,
    #demographics
    GENDER_IDENTITY,
    RACE_RECODE_CAT5,
    SEXMINID,
    AGE,
    #trans-specific variables
    INTERNALIZED_I,
    EVERYDAY_I,
    CHILDGNC_I,
    NONDISCLOSURE_I,
    NEGEXPFUTURE_I,
    SOCSUPPORT_I,
    CONNECTEDNESS_I,
    NONAFFIRM_I,
    HCTHREAT_I,
    #positive health outcomes
    LIFESAT,
    SOCIALWB
  )
```

```{r fitting trees, echo=TRUE, message=FALSE, warning=FALSE}
# Life satisfaction - demographics only
lifesat_tree1 = ctree(
  formula = as.ordered(LIFESAT) ~ AGE + GENDER_IDENTITY + RACE_RECODE_CAT5 +
    SEXMINID,
  data = transonly,
  minbucket = 26,
  multiway = T,
  na.action = na.omit,
  weights = WEIGHT_TRANSPOP
)

# Social wellbeing - demographics only
socialwb_tree1 = ctree(
  formula = as.ordered(SOCIALWB) ~ AGE + GENDER_IDENTITY + RACE_RECODE_CAT5 +
    SEXMINID,
  data = transonly,
  minbucket = 26,
  multiway = T,
  na.action = na.omit,
  weights = WEIGHT_TRANSPOP
)

# Life satisfaction - stressors only
lifesat_tree_stressors = ctree(
  formula = as.ordered(LIFESAT) ~ INTERNALIZED_I +
    EVERYDAY_I + CHILDGNC_I + NONDISCLOSURE_I +
    NEGEXPFUTURE_I + SOCSUPPORT_I + CONNECTEDNESS_I +
    NONAFFIRM_I + HCTHREAT_I + SEXMINID,
  data = transonly,
  minbucket = 26,
  multiway = T,
  na.action = na.omit,
  weights = WEIGHT_TRANSPOP
)

# social wellbeing - stressors only
socialwb_tree_stressors = ctree(
  formula = as.ordered(SOCIALWB) ~ INTERNALIZED_I +
    EVERYDAY_I + CHILDGNC_I + NONDISCLOSURE_I +
    NEGEXPFUTURE_I + SOCSUPPORT_I + CONNECTEDNESS_I +
    NONAFFIRM_I + HCTHREAT_I + SEXMINID,
  data = transonly,
  minbucket = 26,
  multiway = T,
  na.action = na.omit,
  weights = WEIGHT_TRANSPOP
)

```

```{r echo=FALSE}
lifesat_tree1
plot(lifesat_tree1)
```

```{r echo=FALSE}
socialwb_tree1
plot(socialwb_tree1)
```

```{r echo=FALSE}
lifesat_tree_stressors
plot(lifesat_tree_stressors)
```

```{r echo=FALSE}
socialwb_tree_stressors
plot(socialwb_tree_stressors)

```

------------------------------------------------------------------------

References

Krueger, E. A., Divsalar, S., Luhur, W., Choi, S. K., & Meyer, I. H. (2020). TransPop - U.S. Transgender Population Health Survey (Methodology and Technical Notes). Los Angeles, CA: The Williams Institute. Retrieved from: <https://www.transpop.org/s/TransPop-Survey-Methods-v18-FINAL-copy.pdf>

Meyer, I. H. (2021). TransPop, United States, 2016-2018 [Dataset]. Inter-university Consortium for Political and Social Research [distributor]. <https://doi.org/10.3886/ICPSR37938.v1>
